Universals of word order result from optimization of grammars for efficient communication

### 2019-03-04 second-draft/3-paper-names.tex
revising

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{MH and RF designed research. MH implemented models and experiments. MH and RF wrote the paper. MH, DJ, and RF provided comments on the paper.}
\authordeclaration{The authors declare no conflict of interest.}
%\equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: mhahn2@stanford.edu}

@@

\begin{abstract}
The universal properties of human languages have been the subject of intense study across disciplines. We report novel computational and corpus-based evidence for the hypothesis that a prominent subset of these universal properties---those related to word order---result from a process of optimization for efficient communication among humans. We develop a probabilistic, differentiable model of word order grammars: the means by which different languages convert underlying hierarchical structures into strings of words. We show how the parameters of a word order grammar can be optimized for efficiency of information transfer, quantified as a tradeoff between incremental predictability and mutual information with latent tree structures. Applying these grammars to tree structures found in dependency corpora from 51 languages, we show that optimizing the grammar parameters for efficiency results in word order patterns that reproduce a large subset of the major word order correlations reported in the linguistic typological literature, and reproduce the predictions of previous heuristic theories such as dependency length minimization.
\end{abstract}

@@

\textbf{TODO maybe should be more general in the beginning, as in the first version. Also make sure the reader doesn't think 'it's just headedness, why bother'.}

Natural languages vary a lot in the order in which they express information.
Consider Figure~\ref{fig:arabic-japanese-simple}, showing a sentence in Arabic (top) and Japanese (bottom), both translating to `I wrote a letter to a friend.'
Both sentences contain a verb meaning `wrote', a noun expressing  `letter', and a phrase translating to `to a friend'.
In linguistic terminology, `letter' is the object of the verb, whereas the phrases translating as `to a friend' are known as adpositional phrases.
While the two sentences containing words with the same meanings, the order of these words are entirely different in the two languages:
%However, the languages differ in the order in which these words and phrases are arranged: 
In Arabic, the verb stands at the beginning, followed by both the object and the adpositional phrase.
`To' is expressed by a \emph{preposition}, so named because it precedes the word expressing `friend'. %The adpositional phrase starts with the adposition.
In Japanese, the verb stands at the end; object and adpositional phrase precede it; `to' is expressed by a \textit{postposition}, so named because it follows the word denoting `friend'.
%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb, and that the objects of adpositions and verbs stand on the sam side of their heads.
%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb.
%It turns out that these are deep and stable regularities:
It turns out that this variation reflects a deep and stable regularity:
While languages ordering the objects before (Japanese) or after (Arabic) the verb are approximately equally common around the world,
this is strongly correlated with the occurrence of pre- or postpositions:
Languages ordering their objects the way Japanese does, have postpositions; languages ordering them as Arabic does have prepositions.
%However, the vast majority of natural languages place adpositional phrases on the same side of the verb as the object:
This correlation has an extremely strong empirical basis:
In a sample of about 1,100 languages from five contents (CITE Dryer WALS 95a), less than 7 \% of languages provide clear exceptions to this generalization. %, over 90 \% of languages satisfy this generalization .

@@

A commonly proposed functional explanation of Greenberg's correlations is in terms of Dependency Length Minimization \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
This is the idea that languages order words so as to minimize the linear distance between words linked in head--dependent relationships.
This has been confirmed on the basis of large-scale corpus evidence \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}, and has been justified based on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}.

@@

Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between speaker effort (Force of Unification) and listener effort (Force of Diversification).
According to this theory, a language that minimizes speaker effort should reduce the number of different utterances to make production as easy as possible.
Conversely, to minimize listener effort, the language should provide different utterances for different meanings, so that the listeenr can unambiguously identify the meaning from the utterance.
Minimizing speaker and listener effort represent two opposing forces:
A pressure to reduce the complexity of the language makes the utterances in a language more homogenous.
A pressure to reduce ambiguity makes utterances more heterogeneous, so they can indicate different meanings.
%Another way to conceptualize these two pressures is as the \key{cost} and \key{informativity} of language use:
Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required to be provide higher informativity.
The idea that language results from the tension between these two pressures has a long and fruitful history.
%\cite{horn-toward-1984}, RSA, Color naming
%
%In the domain of word order, 
%
%TODO introduce the terms \key{cost} and \key{informativity}
%
This idea has been shown to account for phenomena such as pragmatic inference \cite{horn-toward-1984, frank2012predicting} and color naming~\cite{zaslavsky2018efficient}.

@@

In this work, we show that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations.

@@

\begin{table*}
	\begin{center}
\begin{tabular}{|ll|l|l|l|ll|l|llllllll}
	\hline
	%&
	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
%adp. &
	adposition    &    NP       \\
	\emph{to}            & \emph{a friend} \\
copula    &    NP        &     \\
	\emph{is}        & \emph{a friend} \\
auxiliary    &    VP       &   \\
	\emph{has}          & \emph{written} \\
noun    &    genitive      &  \\
	\emph{friend} &  \emph{of John} \\
noun    &    relative clause       &  \\
	\emph{books} & \emph{that you read} \\
complementizer    &    S        & \\
	\emph{that} & \emph{Mary} \\
verb    &    PP     &    \\
	\emph{went} & \emph{to school} \\
want    &    VP      &    \\
	\emph{wants}   &  \emph{to leave} \\
verb    &    subject    &  \\
	\emph{(there) entered} & \emph{a person} \\
verb    &    manner adverb     & \\
	\emph{ran} & \emph{quickly} \\
 \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
\end{tabular}
	\end{center}
	\caption{Greenbergian Correlations. Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
For each correlation, we provide an example.
Given the statistical nature of the correlations, not every language satisfies every one of them.
Not all correlations are satisfied by every natu
}\label{table:corr-dryer}
\end{table*}

@@

All natural languages have some degree of word order regularities that are specified in their grammar.
For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.
That is, speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences.
Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
%Speakers encode dependency trees into sentences according to a grammar~(Figure~\ref{fig:grammars}).
%While natural languages differ in the freedom of word order, all languages rely on some amount of word order reglarity to encode meaning.
This is illustrated in Figure~\ref{fig:grammars}: Grammars specify how dependency structures are encoded into strings of words.
For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.
%In a speech community, speakers and listeners have shared knowledge of these rules, which allows the listener to decode the dependency tree.
%In the terminology of information theory, these rules specify a \emph{channel} through which the speaker transmits the dependency structures.

@@

@@

>>>>

% Please include corresponding author, author contribution and author declaration information
%\authorcontributions{MH and RF designed research. MH implemented models and experiments. MH and RF wrote the paper. MH, DJ, and RF provided comments on the paper.}
\authordeclaration{The authors declare no conflict of interest.}
%\equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: mhahn2@stanford.edu}

@@

\begin{abstract}
The universal properties of human languages have been the subject of intense study across disciplines. We report novel computational and corpus-based evidence for the hypothesis that a prominent subset of these universal properties---those related to word order---result from a process of optimization for efficient communication among humans. We develop a probabilistic, differentiable model of word order grammars: the means by which different languages convert underlying hierarchical structures into strings of words. We show how the parameters of a word order grammar can be optimized for efficiency of information transfer, quantified as a tradeoff between incremental predictability and mutual information with latent tree structures. Applying these grammars to tree structures found in dependency corpora from 51 languages, we show that optimizing the grammar parameters for efficiency results in word order patterns that predict a large subset of the major word order correlations reported in the linguistic typological literature, and subsumes the predictions of previous heuristic theories such as dependency length minimization.
\end{abstract}

@@

%\textbf{TODO maybe should be more general in the beginning, as in the first version. Also make sure the reader doesn't think 'it's just headedness, why bother'.}

Natural languages vary a lot in the order in which they express information.
Consider Figure~\ref{fig:arabic-japanese-simple}, showing a sentence in Arabic (top) and Japanese (bottom), both translating to `I wrote a letter to a friend.'
Both sentences contain a verb meaning `wrote', a noun expressing  `letter', and a phrase translating to `to a friend'.
In linguistic terminology, `letter' is the object of the verb, whereas the phrases translating as `to a friend' are known as adpositional phrases.
While the two sentences containing words with the same meanings, the order of these words are entirely different in the two languages:
%However, the languages differ in the order in which these words and phrases are arranged: 
In Arabic, the verb stands at the beginning, followed by both the object and the adpositional phrase.
`To' is expressed by a \emph{preposition}, so named because it precedes the word expressing `friend'. %The adpositional phrase starts with the adposition.
In Japanese, the verb stands at the end; object and adpositional phrase precede it; `to' is expressed by a \textit{postposition}, so named because it follows the word denoting `friend'.
%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb, and that the objects of adpositions and verbs stand on the sam side of their heads.
%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb.
%It turns out that these are deep and stable regularities:
It turns out that this variation reflects a deep and stable regularity:
While languages ordering the objects before (Japanese) or after (Arabic) the verb are approximately equally common around the world,
this is strongly correlated with the occurrence of pre- or postpositions:
This is illustrated in Figure~\ref{fig:corr-table}.
Languages ordering their objects the way Japanese does, have postpositions; languages ordering them as Arabic does have prepositions.
%However, the vast majority of natural languages place adpositional phrases on the same side of the verb as the object:
As Figure~\ref{fig:corr-table} shows, this correlation has an extremely strong empirical basis:
In a sample of about 1,100 languages from five contents (CITE Dryer WALS 95a), less than 7 \% of languages provide clear exceptions to this generalization. %, over 90 \% of languages satisfy this generalization .

\begin{figure}
\centering
\begin{dependency}[theme = simple]
   \begin{deptext}[column sep=1em]
%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
	   katabt \& risāla \& li \& sadīq  \\
%	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
	   wrote \& letter   \& to \& friend  \\
   \end{deptext}
\end{dependency}
\begin{dependency}[theme = simple]
   \begin{deptext}[column sep=1em]
      tomodachi \& ni \& tegami-o \& kaita \\
 %\textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
	   friend \& to \& letter  \& wrote \\
   \end{deptext}
\end{dependency}

    \centering
    \includegraphics[width=0.4\textwidth]{figures/corr-table.png}

%\begin{dependency}[theme=simple]
%    \begin{deptext}
%    I \& wrote \& a \& letter \& to \& a \& friend \\
%    \end{deptext}
%    \depedge{2}{1}{nsubj}
%    \depedge{2}{4}{obj}
%    \depedge{4}{3}{ }
%    \depedge{4}{3}{ }
%    \depedge{2}{5}{}
%    \depedge{5}{7}{ }
%    \depedge{7}{6}{ }
%\end{dependency}
	\caption{ The idea of a word order correlation: Languages can order the object before or after the verb (rows), and have prepositions or postpositions (columns). For each combination, we indicate how many languages satisfy it, as documented in the World Atlas of Language Structure~\cite{wals}. Combinations on the diagonal are vastly more common than off-diagonal ones.  A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.}\label{fig:arabic-japanese-simple}	\label{fig:corr-table}
\end{figure}

@@

Theoretical studies of word order typology have proposed Dependency Length Minimization as a unifying principle behind Greenberg's correlations \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
This is the idea that languages order words so as to minimize the linear distance between words linked in head--dependent relationships.
%Our first contribution is to show that this principle indeed correctly predicts Greenberg's correlations.
%We do this by creating counterfactual grammars that minimize dependency length of 
This principle is a heuristic formalization of the idea that long dependencies should create high memory requirements in language processing \cite{gibson1998linguistic,gibson2000dependency}.
It has been argued theoretically that this principle should account for Greenberg's correlations, though this link is nontrivial and depends on the precise distributions of tree topologies in syntactic structures.

%We have verified that dependency length minimization produces Greenberg's correlation universals.
%However, dependency length is only a heuristic and stipulative measure of complexity: % heuristic and stipulative.
%The actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}. 
%It is also stipulative: There is no a priori reason why dependency length should be the crucial determining factor determining word order.
%While DLM has been justified based on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}, there are other models of processing cost that make different predictions.
%In particular, surprisal is known to be a strong predictor of human processing effort, but it makes predictions different from and sometimes opposite to those made by dependency length.
%DLM is also insufficient as an underlying principle of word order, as absolutely minimizing dependency length results in languages that have no consistent grammatical orders, making comprehension difficult.
%Indeed, theoretical work has justified DLM as describing an aspect of a broader notion of language efficiency.
%\cite{hawkins1994performance} theoetically argues that DLM enables more efficient parsing.
%\cite{futrell2017noisycontext} argues that DLM arises from the optimization of predictability.

%A commonly proposed functional explanation of Greenberg's correlations is in terms of Dependency Length Minimization \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
%This is the idea that languages order words so as to minimize the linear distance between words linked in head--dependent relationships.
%This has been confirmed on the basis of large-scale corpus evidence \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}, and has been justified based on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}.

@@

Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between speaker effort (Force of Unification) and listener effort (Force of Diversification).
According to this theory, a language that minimizes speaker effort should reduce the number of different utterances to make production as easy as possible.
Conversely, to minimize listener effort, the language should provide different utterances for different meanings, so that the listeenr can unambiguously identify the meaning from the utterance.
Minimizing speaker and listener effort represent two opposing forces:
A pressure to reduce the complexity of the language makes the utterances in a language more homogenous.
A pressure to reduce ambiguity makes utterances more heterogeneous, so they can indicate different meanings.
%Another way to conceptualize these two pressures is as the \key{cost} and \key{informativity} of language use:
Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required to be provide higher informativity.
The idea that language results from the tension between these two pressures has a long and fruitful history.
%\cite{horn-toward-1984}, RSA, Color naming
%
%In the domain of word order, 
%
%TODO introduce the terms \key{cost} and \key{informativity}
%
This idea has been shown to provide a formalization and derivation of pragmatic inference \cite{horn-toward-1984, frank2012predicting}.
It also has been shown to account for the distribution of word meanings across languages, such as color naming~\cite{zaslavsky2018efficient}.

@@

In this work, we show that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations, and for Dependency Length Minimization.

@@

\begin{table*}
	\begin{center}
\begin{tabular}{|ll|l|l|l|ll|l|llllllll}
	\hline
	%&
	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
%adp. &
	adposition    &    NP       
	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf} \textbf{86}  }
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_case.pdf}   \textbf{81}$^{***}$    }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf}   \textbf{47 \%}$^{}$   }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_case.pdf}  \textbf{76}$^{***}$ }   
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf} \textbf{68 \%}$^{***}$    } \\
	\emph{to}            & \emph{a friend} \\ \hline
copula    &    NP         
	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf} \textbf{94} }
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_cop.pdf}   \textbf{81}$^{***}$    }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf}  \textbf{53}     }
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_cop.pdf}  \textbf{76}$^{***}$ }   
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf} \textbf{61}$^{***}$  }  \\
	\emph{is}        & \emph{a friend} \\ \hline
auxiliary    &    VP       
	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf}}
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-aux.pdf}   \textbf{81}$^{***}$    }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf}   }
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-aux.pdf}   \textbf{76}$^{***}$ }   
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}}  \\
	\emph{has}          & \emph{written} \\ \hline
noun    &    genitive      
	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf}}
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf} \textbf{81}$^{***}$    }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf}   }
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf} \textbf{76}$^{***}$ }   
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}}  \\
	\emph{friend} &  \emph{of John} \\ \hline
noun    &    relative clause      
	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf}}
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-acl.pdf}\textbf{81}$^{***}$    }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf}   }
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-acl.pdf} \textbf{76}$^{***}$ }   
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}}  \\
	\emph{books} & \emph{that you read} \\ \hline
complementizer    &    S        
	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf}}
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_mark.pdf} \textbf{81}$^{***}$    }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf}   }
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_mark.pdf}\textbf{76}$^{***}$ }   
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}}  \\
	\emph{that} & \emph{Mary has arrived} \\ \hline
verb    &    PP         
	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf}}
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-obl.pdf} \textbf{81}$^{***}$    }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf}   }
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-obl.pdf}\textbf{76}$^{***}$ }   
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}}  \\
	\emph{went} & \emph{to school} \\ \hline
want    &    VP        
	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf}}
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf} \textbf{81}$^{***}$    }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf}   }
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf} \textbf{76}$^{***}$ }   
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf}}  \\
	\emph{wants}   &  \emph{to leave} \\ \hline
 \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
\end{tabular}
	\end{center}
	\caption{Greenbergian Correlations. Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
For each correlation, we provide an example.
Given the statistical nature of the correlations, not every language satisfies every one of them.
Not all correlations are satisfied by every natu
}\label{table:corr-dryer}
\end{table*}

@@

All natural languages have some degree of word order regularities that are specified in their grammar.
For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.
That is, speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences.
Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
%Speakers encode dependency trees into sentences according to a grammar~(Figure~\ref{fig:grammars}).
%While natural languages differ in the freedom of word order, all languages rely on some amount of word order reglarity to encode meaning.
This is illustrated in Figure~\ref{fig:grammars}: Grammars specify how dependency structures are encoded into strings of words.
For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.
%In a speech community, speakers and listeners have shared knowledge of these rules, which allows the listener to decode the dependency tree.
%In the terminology of information theory, these rules specify a \emph{channel} through which the speaker transmits the dependency structures.
We formalize word order grammars as specifying how different types of dependents (subjects, objects, etc.) are ordered, employing a cross-linguistically meaningful inventory of relations (CITE UD).

@@

Discuss DLM now again

@@

\subsection{Greenbergian Word Order Correlations}
\label{sec:greenberg}

%\begin{table*}
%	\begin{center}
%\begin{tabular}{|ll|l|l|l|ll|l|}
%	\hline
%	%&
%	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
%	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
%%adp. &
%adposition    &    NP       &    86    &    \textbf{81}$^{***}$    &    47    &    \textbf{76}$^{***}$    &    \textbf{68}$^{***}$   \\
%copula    &    NP        &    94    &    \textbf{81}$^{***}$    &    53    &    \textbf{79}$^{***}$    &    \textbf{61}$^{**}$   \\
%auxiliary    &    VP       &    88    &    \textbf{74}$^{***}$    &    \textbf{84}$^{***}$    &    55    &    \textbf{69}$^{**}$   \\
%noun    &    genitive      &    80    &    \textbf{82}$^{***}$    &    55    &    \textbf{74}$^{***}$    &    \textbf{70}$^{***}$   \\
%noun    &    relative clause       &    80    &    \textbf{85}$^{***}$    &    48    &    \textbf{77}$^{**}$    &    \textbf{73}$^{***}$   \\
%complementizer    &    S        &    76    &    \textbf{85}$^{***}$    &    \textbf{59}$^{**}$    &    \textbf{80}$^{***}$    &    \textbf{74}$^{**}$   \\
%verb    &    PP     &    88    &    \textbf{78}$^{***}$    &    \textbf{72}$^{***}$    &    59    &    \textbf{69}$^{**}$   \\
%want    &    VP      &    88    &    \textbf{90}$^{***}$    &    \textbf{78}$^{**}$    &    \textbf{92}$^{***}$    &    \textbf{92}$^{***}$   \\
%verb    &    subject    &    33    &    \textbf{29}$^{**}$    &    51    &    \textbf{8}$^{***}$    &    \textbf{13}$^{***}$   \\
%verb    &    manner adverb     &    35    &    51    &    \textbf{21}$^{***}$    &    51    &    \textbf{32}$^{***}$   \\
% \hline
%%comp. &
%%\multirow{3}{*}{NP}&
%%	AP &
%    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
%\end{tabular}
%	\end{center}
%	\caption{Greenbergian Correlations. Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
%%	For clarity, we organize the correlations by the category of the `verb patterner'.
%	For each correlation, we give our operationalization in terms of UD. %A rightwards arrow indicates that the dependency is predicted to have the same direction as the \textit{obj} dependency; a leftwards arrow indicates an inverse correlation. 
%	For each correlation we report what percentage of the languages in our sample satisfied it (`Real').
%	We then report, for each correlation and each objective function, how many (in \%) of the optimized grammars satisfy the correlation, with the significance level in a logistic mixed-effects analysis across language families.}\label{table:results-dryer}
%\end{table*}
%
